---
title: "Text Mining Homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, fig.align = "center", message = FALSE, warning = FALSE)
```

# MVP

Using the dataset `austen_books()` from the package `janeaustenr`:

1. Find the most common words in both Pride & Prejudice and Sense & Sensibility.
2. Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.
3. Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

```{r, message = FALSE}
library(dplyr)
library(tidytext)
library(janeaustenr)
```

```{r}
# 1 
austen_words <- 
austen_books() %>%
  filter(book == "Pride & Prejudice" | book == "Sense & Sensibility") %>%
  unnest_tokens(word, text) %>%
  count(word, book, sort = TRUE)
```
```{r}
austen_words %>%
  filter(book == "Pride & Prejudice")
```
```{r}
austen_words %>%
  filter(book == "Sense & Sensibility")
```

```{r}
# 2
austen_words <- 
austen_words %>%
  anti_join(stop_words)
```

```{r}
austen_words %>%
  filter(book == "Pride & Prejudice")
```
```{r}
austen_words %>%
  filter(book == "Sense & Sensibility")
```

```{r}
# 3
austen_sentiment_words <-
  austen_words %>%
  inner_join(get_sentiments("bing"))
```

```{r}
austen_sentiment_words %>%
  filter(book == "Pride & Prejudice")
```

```{r}
austen_sentiment_words %>%
  filter(book == "Sense & Sensibility")
```

# Extension

Taking your results above. Can you create a plot which visualises the differences between the books?  

```{r}
library(ggplot2)

austen_words <-
austen_words %>%
mutate(
  sentiment = "Neutral"
)
  
plot_df <- 
rbind(
  slice(austen_words, 1:10),
  slice(austen_sentiment_words, 1:10)
)

ggplot(plot_df) +
  aes(x = word, y = n, fill = book) +
  geom_col(position = "dodge") +
  coord_flip() +
  facet_wrap(sentiment ~ ., scales = "free")
```




