---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(modelr)
```

```{r}
avocado_data <- read_csv("data/avocado.csv")
avocado_data
```
```{r}
summary(avocado_data)
```

```{r}
glimpse(avocado_data)
```

#looks like total volume includes PLU other than hass classification therefore deselect hass and focus on total volume sold as looking at average_price
```{r}
avocado_data <- avocado_data %>% 
  clean_names() 
  
avocado_data
```
```{r}
glimpse(avocado_data)
```


```{r}
tidy_avocado_data <- avocado_data %>% 
  #maybe extract month from date before deselection to identify seasonal rates of production? 
    select(-c("region", contains("bags"), year, date)) %>% 
      mutate(type = as.factor(type)) %>% 
        rename( total_number_sold = total_volume,
            average_price = average_price,
              small_hass = x4046,
              medium_hass = x4225,
              large_hass = x4770) 
             # select(-c(contains("hass")))
tidy_avocado_data
```

#build model systematically
```{r}
ggpairs(tidy_avocado_data)
```
#average price has some correlation with small hass, total number and type
```{r}
tidy_avocado_data %>%
  ggplot(aes(x = type, y = average_price, group = 1)) +
  geom_boxplot()
```

```{r}
tidy_avocado_data %>%
  ggplot(aes(x = small_hass, y = average_price, group = 1)) +
  geom_boxplot()
```

```{r}
tidy_avocado_data %>%
  ggplot(aes(x = total_number_sold, y = average_price, group = 1)) +
  geom_boxplot()
```



```{r}
mod1a <- lm(average_price ~ type, data = tidy_avocado_data)
summary(mod1a)

mod1b <- lm(average_price ~ small_hass, data = tidy_avocado_data)
summary(mod1b)

mod1c <- lm(average_price ~ total_number_sold, data = tidy_avocado_data)
summary(mod1c)
```

#r^2  is lowish (0.38) can explain 38% of the variance in average_price based on type. 

```{r}

par(mfrow = c(2, 2)) 
plot(mod1a)

par(mfrow = c(2, 2)) 
plot(mod1b)

par(mfrow = c(2, 2)) 
plot(mod1c)
```

now plot diagnostics

#spent too long on working through weeks notes and didn't leave enough time to work through hw. again!

