---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
library(tsibble)
library(tsibbledata)
library(feasts)

date1 <- ymd_hms("2017-04-22 22:01:18")
date2 <- ymd_hms("2001-11-07 10:09:56")
```

```{r}
library(nycflights13)
```


**Question 1**  

Extract the different components (year, month, mday, yday, wday) from the variable `date1`. Add the labels to the month and do not abbreviate the month labels. 


```{r}
as_datetime(date1)
year(date1)
month(date1, label = TRUE, abbr =FALSE)
mday(date1)
yday(date1)
wday(date1)

```

**Question 2**  

Add 14 days to your `date1` variable and store it in a variable called `next_fortnight`. 
Take away 10 years from your `date2` variable, and store it in `previous_decade`. 

```{r}
next_fortnight <- date1 + days(14)
next_fortnight
```

```{r}
previous_decade <- date2 - years(10)
previous_decade
```

**Question 3**  

Create a variable called `this_year` which contains today's date. Then create an interval object which uses `this_year` and the `previous_decade` variable you made above. Store it in `interval_years`. 
*Hint: you'll need to use lubridate::interval in order to get the right function, as tsibble has one too*  


```{r}
this_year <- today()
this_year

interval_years <- lubridate::interval(this_year, previous_decade)
interval_years
```



**Question 4**    

Change the time zone of both your date variables to "America/New_York", and call them `nyc_date1` and `nyc_date2`. 

```{r}
nyc_date1 <- with_tz(date1, tzone = "America/New_York")
                     
nyc_date1

  
nyc_date2 <- with_tz(date2, tzone = "America/New_York")

nyc_date2

```


**Question 5**  

Use the `flights` data from the `nycflights13` dataset. Make a date-time column called `departure_date` from the `year, month, day, hour, and minute` variables in the `flights` tibble.

```{r}
nycflights13::flights
```

```{r}
nycflights13::flights %>% 
  mutate(departure_date = make_datetime(year, month, day, hour, minute))

```


# Time series familiarity 


**Question 6**  

Start with the `weather` data from the `nycflights13` package. Select all the variables related to temperature, humidity, and precipitation, and pick out the key and index variables. Have a look at your data, and write a small description of the time series you observe. 

Here we can see that our data is collected in hourly format, with observations occuring every day. 
```{r}

library(tsibbledata)
data(package = "tsibbledata")
```

```{r}
library(tsibble)
```

```{r}
nycflights13::weather
```

```{r}
nyc_weather <- nycflights13::weather %>% 
select(origin, temp, humid, precip, time_hour)
nyc_weather
```

```{r}
class(nyc_weather)
#tsibble(nyc_weather)
```


key-? origin, after I read below.
index- time component of data: time_hour


**Question 7**  

Change the weather data into a tsibble. Hint, look up in the notes where we suggest using `as_tsibble` function. You'll need to define an `index` and `key`.


```{r}
as_tsibble(nyc_weather, key = origin, index = time_hour)
```


In order to create a time series tibble, remember you need to provide an *index* and a *key*. In our case, `time_hour` is the only column involving a date/date-time, and so this would be our index. A `tsibble` also requires a key, which defines individual observations that are tracked over time. In this example, the `origin` variable is the identifier, and therefore is our key. It is important to note that the duo of the index and key uniquely identifies each observation in a valid `tsibble`.   

**Question 8**  

Add a year and month column to your new tsibble weather data, 

```{r}
nyc_weather <- nyc_weather %>% 
    mutate(year = year(time_hour), month = month(time_hour))

nyc_weather
```

and create a basic line plot which shows the temperature data from each airport over time.

```{r}
nyc_weather  %>% 
  select(origin, month, temp) %>% 
  group_by(origin, month)  %>% 
  summarise(mean_temp = mean(temp))
```
```{r}
nyc_weather %>% 
  ggplot() +
  geom_line(aes(x = time_hour, y = temp, colour = origin))
```


Have a think back to what you learned in visualisation week, and think about whether plotting the whole time series in one go is adequate here? 

##No. shows hour by hour variation where daily or monthly would be sufficient.

Write down your suggestions for what could be done to make a better and more informative plot. 

##select data over a more finite time series, ie over days or months as opposed to over hours.

```{r}
nyc_weather %>% 
  ggplot() +
  geom_line(aes(x = month, y = temp, colour = origin))
```


**Question 9**  

Choose an appropriate rolling window and calculate a moving average for your weather temperature data for each airport origin. Create a useful plot which overlays the original time series data and the rolling average data. Make sure the graph is easy for others to understand (*Hint: you might want to facet_wrap*). Finally, write a summary of why you chose the time window you did, and describe any trends you see. 



**Question 10**  

Downsample your data so that you plot daily temperature data for each airport, instead of hourly data. 


<br>

