---
title: "R Notebook"
output: html_notebook
---

* You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings.

* We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

```{r}
library(tidyverse)
library(janitor)
```

```{r}
customers <- read_csv("clustering_homework/mall_customers.csv")

customers
```

```{r}
customers <-customers %>% 
  clean_names()
```

 
* Perform k-means clustering and chose a value of k.
#select on variables to be investigated in method 
```{r}
customers <- customers  %>% 
              select(c(annual_income_k, spending_score_1_100))
summary(customers)
```


plot to look for clustering: ##looks like 5 potential clusters
```{r}
customers %>% 
ggplot(aes(annual_income_k, spending_score_1_100)) +
  geom_point()
```

#scale
```{r}
customers_scale <- customers %>%
                      mutate_all(scale)

summary(customers_scale)
```

#k-means calculation. k = 5, n =10(more than  but not so much to impact computing time)
```{r}
clustered_customers <- kmeans(customers_scale, centers = 5, nstart = 10)
clustered_customers
```



```{r}
library(broom)

tidy(clustered_customers, 
     col.names = colnames(customers_scale))
```

```{r}
glance(clustered_customers)
```


```{r}
augment(clustered_customers, customers)
```

elbow method- weak suggestion of k =5. investigate further
```{r}

# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters_customers <- tibble(k = 1:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(customers_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customers)
  )

k_clusters_customers
```

```{r}
clusterings <- k_clusters_customers%>%
  unnest(glanced)

clusterings
```

plot results
```{r}
ggplot(clusterings, aes(x=k, y=tot.withinss)) +
  geom_point() +
    geom_line() +
    scale_x_continuous(breaks = seq(1, 20, by = 1))
```

quicker method for all 3
```{r}
library(factoextra)

#elbow
fviz_nbclust(customers_scale, kmeans, method = "wss", nstart = 25)
```

#silhouette suggests k =5
```{r}
#silhouetter
fviz_nbclust(customers_scale, kmeans, method = "silhouette", nstart = 25)
```

#gap stat also suggests k = 5
```{r}
#gap statistic
fviz_nbclust(customers_scale, kmeans, method = "gap_stat", nstart = 25, k.max = 10)
```


* Visualise the clustering for your chosen value of k.
#visualise k value. k =7? but previous ,ethods support 5
```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k <= 10) %>%
 ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point(aes(color = .cluster)) + 
  facet_wrap(~ k)
```

```{r}
 clusterings %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) %>%
 ggplot(aes(x = annual_income_k, 
            y = spending_score_1_100, 
            colour = .cluster, 
            label = .rowna)) +
   geom_point(aes(color = .cluster)) #+
  #geom_text(hjust = 0, vjust = - 0.5, size = 3)
```


* Do you think the clustering seems a good fit for this data?
Yes

* Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 

#cluster 1- mid income 
