---
title: "Clustering weekend homework - solutions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
library(janitor)
```


```{r}
customers <- read_csv("mall_customers.csv") %>%
                clean_names()
```

Do some exploratory analysis on the data. Comment on findings. 

```{r}
summary(customers)
```

```{r}
customers %>%
  group_by(gender) %>%
  summarise(count = n()) %>%
ggplot(aes(x = gender, y = count)) +
  geom_col()
```

```{r}
ggplot(customers, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()
```

```{r}
ggplot(customers, aes(x = annual_income_k)) +
  geom_histogram()
```


```{r}
ggplot(customers, aes(x = spending_score_1_100)) +
  geom_histogram()
```


```{r}
ggplot(customers, aes(x = age)) +
  geom_histogram(bins = 50)
```


We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers. 

```{r}
customers_scale <- customers %>% 
  select(annual_income_k, spending_score_1_100) %>%
    mutate_all(scale)

summary(customers_scale)
```

```{r}
ggplot(customers_scale, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()
```

There looks to be dispersion in the data and can see some potential cluster groupings, so clustering looks like it may be appropriate here. 

```{r}
library(broom)

# Set min & max number of clusters want to look at 
max_k <- 20 

k_clusters <- tibble(k = 2:max_k) %>%
  mutate(
    kclust = map(k, ~ kmeans(customers_scale, .x, nstart = 25)), 
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, customers_scale)
  )

k_clusters
```


```{r}
library(factoextra)
fviz_nbclust(customers_scale, kmeans, method = "wss", nstart = 25)
fviz_nbclust(customers_scale, kmeans, method = "silhouette", nstart = 25)
fviz_nbclust(customers_scale, kmeans, method = "gap_stat", nstart = 25)
```


There looks to be a clear defined elbow at k = 5 and this was also selected for the silhouette method so go ahead with k = 5. 

```{r}
customer_clusters_k5 <- k_clusters %>% 
  unnest(cols = c(augmented)) %>%
  filter(k == 5) 
```


And let's plot the data:
```{r}
ggplot(customer_clusters_k5, aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point(aes(color = .cluster))
```

Do you think the clustering seems a good fit for this data?
```{r}
#yes based on the clearly defined clusters from the visuals and that there is a majority of a value of k for the 3 methods for selecting k. 
```


Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the 'Segmentation & clustering intro' lesson). 

```{r}
#Cluster number 5 is high income and high spend - these may be more 'high end' customers and could look to target them accordingly. 
#Cluster number 2 is high income but low spend - perhaps they mainly shop at other malls or spend money on activities/items not sold within your mall. Perhaps may want to target with further data collection to find out more about this in case we are able to target with a particular campaign if we find they are shopping at competitors. 
```




