---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse
        )
```

#  describe/summarise at least 5 things about this dataset -Be prepared to share one of your findings on Monday!


```{r}
kaggle_goodreads_books <- read_csv('book_data/books.csv')
kaggle_goodreads_books
```
8472 rows with 12 columns
```{r}
dim(kaggle_goodreads_books)
```

```{r}
names(kaggle_goodreads_books)
```


```{r}
kaggle_goodreads_books %>% 
  arrange(desc(ratings_count))
```



```{r}
kaggle_goodreads_books %>% 
  arrange(desc(text_reviews_count))
```

```{r}
names(kaggle_goodreads_books)
```
group languages
```{r}


```
# group languages and order descending
# next step would be to rename all variations of English to get total value 
# then to invesitgate within this English group as with other language groups

```{r}
kaggle_goodreads_books %>% 
  filter(!is.na(language_code)) %>% 
  group_by(language_code) %>% 
  summarise(
    count = n()
  ) %>% 
  arrange(desc(count)) %>% 
  #trying to change en-US to eng
  mutate(language_code, "en-US" = "eng")
```
#then aggregate all english variations to get total English
#then investigate top English author via average_rating, ratings_count, text_reviews_count


#group by author
#Stephen King has most mentions
```{r}
kaggle_goodreads_books %>% 
  filter(!is.na(authors)) %>% 
  group_by(authors) %>% 
  summarise(
    count = n()
  ) %>% 
  arrange(desc(count))
```




