---
title: "Extra programming tasks"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# For loops & apply 

## Question 1
Create vectors `x` and `y` using the commands 

```{r}
x <- c(1, 2, 9)
y <- c(2, 6, 4)
```

Write a for `loop` to compute the element-wise product vector, saving your results in a new vector z,  $z = x \times y$ (e.g. $z_1 = x_1 \times y_1, \; z_2 = x_2 \times y_2, ...$).  Compare this to `x * y` in R. 


## Question 2

Define `x` to be a sample of size 10 from the N(0, 1) distribution using `x <- rnorm(10)` (don't worry if you've not come across a distribution before, we will be covering probability distributions in statistics, but for now just need to know it generates some random numbers between -1 and 1).  

Replace all values that are less than 0 with 0 using the following methods:  

  a.  the `ifelse()` function
  b.  a `for` loop including an `if` statement
  c.  writing a function (to convert a number to zero if negative) and then use that function in an `apply()` statement
  

<br>

## Question 3

Function `cumsum()` computes the cumulative sum of a vector. Use a `for` loop instead of this built in function to compute the cumulative sum of `x` from exercise 2. Compare your loop to the `cumsum` function. 

<br>

## Question 4

Consider the following lines of code that work on a matrix `x` containing observations of several real-valued variables:

```{r, eval=FALSE}
x.min <- apply(x, 2, min)
x.max <- apply(x, 2, max) 

for (i in 1:nrow(x)) {
  x[i, ] <- (x[i, ] - x.min) / (x.max - x.min)
}
```

  a. **Without running the code** explain what each line of the code does. 
  b. **Without running the code** what is the range of the columns of `x` once the code has been run?
  c. Can you rewrite this code using 'loop-avoidance' functions (e.g. `apply()`)? 
  

<br>

## Question 5

Load the library `CodeClanData`, check the help file for the `qb_monthly_sales` dataset and inspect some of the data 

  a. Find the date of the highest costs using the function `which.max()`. 
  b. Find the date of the lowest revenue using the function `which.min()`. 

<br>

# Functions

## Question 6

Write an R function `quadratic()` which computes the real-valued solutions of a quadratic equation

$$a \times x^2 + b \times x + c$$
Your function should take `a`, `b` and `c` as arguments and return the unique solutions to the above quadratic equation (we'll show you the maths below to do this). 

* Your function should start by computing the *discriminant* $\Delta = b^2 - 4 \times a \times c$
* If $\Delta < 0$, then the quadartic has no real solutions. In this case, the function should return a message "No solution".
* If $\Delta = 0$, then the quadratic equation has exactly one ("double") solution, $-\frac{b}{2 \times a}$. In this case, return this value only. 
* If$\Delta > 0$, then the quadratic equation has two solutions, $-\frac{b + \sqrt{\Delta}}{2 \times a}$ and $\frac{b + \sqrt{\Delta}}{2 \times a}$. In this case return a vector of length two containing both solutions. 

<br>

## Question 7

Consider the following function 

```{r}
my_function <- function(z) {
 m <- Inf
 for (i in 1:length(z)) {
   if (z[i] < m) { 
     m <- z[i] 
   }
 }
 return(m)
}
```

**Without running the function** what does `my_function()` do? Can you explain it in plain language?

<br>



# Problem solving

## Question 8

Create the following vector without using the function `c()` at all

    5 6 7 4 5 6 3 4 5 2 3 4
    
<details>
<summary>**Hint**</summary>
Consider creating 2 vectors and adding them, and remember that you can create vectors using other functions such as `rep()` or `seq()`.

</details>
</blockquote>
