---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


```{r}

```

#1_Take the data in the dataset qb_revenue_breakdown and make a stacked bar chart showing the sources of revenue across the two years in the dataset.

```{r}
revenue_breakdown <- CodeClanData::qb_revenue_breakdown
revenue_breakdown
```

```{r}

ggplot(revenue_breakdown)+
  aes(x = Year, y = Revenue, fill = Product) +
geom_col()


```

#2_Make a line chart showing monthly sales of the “flagship product” - the KwikBit Unit1 - over the last year. This data is inside qb_monthly_sales, but you will need to use subset.
```{r}
monthly_sales <- CodeClanData::qb_monthly_sales
monthly_sales
```

```{r}
kwikbit_monthly_sales <- monthly_sales %>% 
    filter(Cashflow == "Kwikbit Unit 1 Sales")
kwikbit_monthly_sales
```


```{r}
ggplot(kwikbit_monthly_sales) +
  aes(x = Date, y = Amount) +
  geom_line() +
  labs(title = "Kwikbit Unit 1 Sales") +
    theme_light()

```


#3_Make a line chart showing monthly revenue and costs over the last year. This data is also in qb_monthly_sales.

```{r}
kwikbit_revenue_and_costs <- monthly_sales %>% 
    filter(Cashflow != "Kwikbit Unit 1 Sales")
kwikbit_revenue_and_costs
```
or

```{r}
revenue_and_costs <- subset(qb_monthly_sales, Cashflow != "Kwikbit Unit 1 Sales")
```



```{r}
ggplot(kwikbit_revenue_and_costs) +
    aes(x = Date, y = Amount, colour = Cashflow) +
    geom_line() +
    labs(title = "Monthly revenue and costs")
    

```


#4Show annual sales of personal fitness trackers over the last 5 years broken down by company as a ribbon plot (use geom_area). This data is in qb_competitors.


```{r}
competitors <- CodeClanData::qb_competitors
competitors
```
```{r}
ggplot(competitors) +
    aes(x = Year, y = Revenue, fill = Company) +
    geom_area() +
    labs(title = "Annual Sales Personal Fitness Trackers")

```


#5Now show the sales from the 5 competitors as a line graph. Include an extra layer that shows the data points used to make the lines.

```{r}
ggplot(competitors) +
    aes(x = Year, y = Revenue, colour = Company) +
    geom_line() +
  geom_point() +
    labs(title = "Annual Sales Personal Fitness Trackers") 
      
```


#6Now the company wants to compare the number of steps that their device counts vs. their competitors. 
#Make a line graph of the number of steps throughout time
#and use faceting to compare between individuals and people. 
#Data is in qb_device_data.


```{r}
device_data <- CodeClanData::qb_device_data
device_data
```

```{r}
names(device_data)
```

```{r}
device_data_mins <- device_data %>% 
  filter(counts != 0) %>% 
  mutate(total_minutes = 60*hours + mins)
device_data_mins
```


```{r}
ggplot(device_data_mins) +
    aes(x = total_minutes, y = counts, colour = device) +
    geom_line() +
    labs(title = "Number of steps through time", y = "step") 
     

```


#have now realised this refers to different data but run out of time! 
```{r}
ggplot(device_data_mins) +
    aes(x = total_minutes, y = counts, colour = device) +
    geom_line() +
    labs(title = "Number of steps through time", y = "step") +
  facet_wrap()



```

