---
title: "R Notebook"
output: html_notebook
---


#1. An .R file that cleans some data on meteorites.
#2. A .Rmd file that takes your cleaned data and performs some analysis.
#3 write a short `README.md` that describes the structure of your project.

#The data you want to clean is in the file called `meteorite_landings.csv`. 

## The .R file

#1. Read the data into R
```{r}
library(tidyverse)
library(janitor)
library(assertr)
```

```{r}
meteorite_landings <- read_csv("../data/raw_data/meteorite_landings.csv")
meteorite_landings
```


#2. Change the names of the variables to follow our naming standards.
```{r}
clean_meteorite_landings <- clean_names(meteorite_landings)
clean_meteorite_landings
```

#3. Split in column `GeoLocation` into latitude and longitude, the new latitude and longitude columns should be numeric.
```{r}
cleaned_meteorite_landings <-clean_meteorite_landings %>% 
    separate(geo_location, c("latitude", "longitude"), 
           sep = ",",convert = TRUE) %>% 
    separate(latitude, c("remove", "latitude"), sep = "\\(")  %>% 
    separate(longitude, c("longitude", "remove_2"), sep = "\\)") %>% 
 
    mutate(latitude, latitude = as.numeric(latitude)) %>% 
    mutate(longitude, longitude = as.numeric(longitude)) %>% 
   
    select(-"remove", -"remove_2") 

cleaned_meteorite_landings

  
```


4. Replace any missing values in latitude and longitude with zeros.
5. Remove meteorites less than 1000g in weight from the data.
6. Order the data by the year of discovery.

```{r}
by_year_cleaned_meteorite_landings <-cleaned_meteorite_landings %>%  
  
    mutate(latitude = coalesce(latitude, latitude, na.rm = TRUE)) %>% 
    mutate(longitude = coalesce(longitude, longitude, na.rm = TRUE)) %>% 

    filter(mass_g > 1000) %>% 
    arrange(year)

by_year_cleaned_meteorite_landings

```



We would also like you to include assertive programming to make sure that:

1. The data has the variable names we expect ("id", "name", "mass (g)", "fall", "year", "GeoLocation").

```{r}
library(assertr)

  names(by_year_cleaned_meteorite_landings)
```

2. Latitude and longitude are valid values. (Latitude between -90 and 90, longitude between -180 and 180).

```{r}
latitude_filter <- filter(by_year_cleaned_meteorite_landings, 
          between(latitude,-90, 90))
latitude_filter
```


```{r}
longitude_latitude_filter <- filter(latitude_filter, 
          between(longitude,-180, 180))  
longitude_latitude_filter
         
```

```{r}
write_csv(longitude_latitude_filter, "tidy_meteorite_landings.csv")
```

```{r}
tidy_meteorite_landings <- read_csv("../data/clean_data/tidy_meteorite_landings.csv")
tidy_meteorite_landings
```

