---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```


```{r}
tidy_meteorite_landings <- read_csv("../data/clean_data/tidy_meteorite_landings.csv")

tidy_meteorite_landings
```

#2. Find the names and years found for the 10 largest meteorites in the data.

```{r}
top_ten_by_mass <- tidy_meteorite_landings %>% 
  arrange(desc(mass_g)) %>% 
  head(n=10) %>% 
  select(name, mass_g, year)

  top_ten_by_mass
```

#3. Find the average mass of meteorites that were recorded falling, vs. those which were just found.

#filter falling versus found
then find mean of each

```{r}
tidy_meteorite_landings %>% 
  filter(fall == "Fell") %>% 
    summarise(mean(mass_g))
              
```

```{r}
tidy_meteorite_landings %>% 
  filter(fall == "Found") %>% 
    summarise(mean(mass_g))
```

example solution
```{r}
tidy_meteorite_landings %>% 
  group_by(fall) %>% 
  summarise(
    avg_mass = mean(mass_g)
  )
```


#4. Find the number of meteorites in each year, for every year since 2000.
group number of meteorites per year since 2000 and summarise

```{r}
tidy_meteorite_landings
```

```{r}
tidy_meteorite_landings %>% 
  group_by(year) %>% 
  count() %>% 
  filter(year >= 2000)

```

example solution
```{r}
tidy_meteorite_landings %>%
  filter(year >= 2000) %>% 
  group_by(year) %>% 
  summarise(n = n())

```


## The project structure

1. Make sure your use an RStudio Project.
2. Make a repo on GitHub and commit your code to this repo regularly.
3. It is up to you how to store your code and data. Make sure it is clear for others using the project.

## The README.md file

Make a plain text file called `README.md`. In this file we just want you to describe the structure of your project, and explain what each piece of code you have written does.