---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
data(msleep)
```

```{r}
glimpse(msleep)
```

#Assume significance level of mu = 5%
inspect distribution
```{r}
msleep %>%
  ggplot(aes(x = sleep_total)) +
  geom_histogram(col = "white")


```

```{r}
msleep %>%
  ggplot(aes(x = sleep_total)) +
  geom_boxplot()
```

#1* Jabberwockies sleep for around $7$ hours a night, on average. Perform an appropriate statistical test to determine whether the mean `sleep_total` in the sampled population of animal types differs from the typical value for jabberwockies.

#Assume significance level α=0.05
Frame question in terms of H0 and Ha.


H0: (the null hypothesis, pronounced ‘H naught’) - the all animal mean_sleep_total for the sampled population is the same as the mean_sleep_total for jabberwockies.

Ha: (the alternative hypothesis) - the all animal mean_sleep_total is different from the jabberwocky mean_sleep_total.

Viewed mathematically:
H0: μmean_sleep_total = 7
Ha: μmean_sleep_total ≠ 7

Observed stat
```{r}
observed_stat <- msleep %>%
  summarise(mean_sleep_total = mean(sleep_total))

observed_stat
```
#Calculate and visualise null sampling distribution

```{r}
library(infer)
```

```{r}
null_distribution <- msleep %>%
  specify(response = sleep_total) %>%
  hypothesize(null = "point", mu = 7) %>%
  generate(reps = 10000, type = "bootstrap") %>%
  calculate(stat = "mean")

```

```{r}
null_distribution %>%
  visualise(bins = 30)
```
Plot observed stat on null distribution
(two tailed test)

```{r}
null_distribution %>%
  visualise(bins = 30) +
  shade_p_value(obs_stat = observed_stat$mean_sleep_total, direction = "both")
```
calculate p value
```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat$mean_sleep_total, direction = "both")
p_value
```

p_value is less than 0.05 therefore reject H0 in favour of Ha.
The mean_sleep_total of 10.43 hours is statistically significantly different to the jabberwocky average of 7hours.

```{r}
glimpse(msleep)
```

#2* Perform an appropriate statistical test to determine whether omnivores sleep for significantly longer than herbivores, on average.

-Samples are independent.
-significance level α=0.05 This will be a one-tailed test:

H0: μsleep_total(omnivores)−μsleep_total(herbivores)=0
Ha: μsleep_total(omnivores)−μsleep_total(herbivores)>0
-group by vore: omni and herbi


```{r}
omni_herbi <- msleep %>% 
       filter(vore ==  "omni" | vore == "herbi") 

omni_herbi
```


summarise and plot distributions
```{r}
omni_herbi %>%
  group_by(vore) %>%
  summarise(n = n())
```


```{r}
omni_herbi %>%
  ggplot(aes(y = sleep_total, x = vore)) +
  geom_boxplot()
```
The omni and herbi samples have a small area of overlap, though the omni range is much bigger.check whether this difference in distributions could be down to sampling variation (i.e. it may have occurred ‘by chance’) or whether it is a statistically significant difference by performing our hypothesis test.

Generate Null distribution by permutation(by creating the sample statistic repeatedly)
```{r}
null_distribution <- omni_herbi %>% 
  specify(sleep_total ~ vore) %>%
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in means", order = c("omni", "herbi")) 

head(null_distribution)
```

calculate our observed statistic and  visualise the null distribution and where the observed statistic lies on the distribution.
```{r}
observed_stat <- omni_herbi %>% 
  specify(sleep_total ~ vore) %>%
  calculate(stat = "diff in means", order = c("omni", "herbi"))

observed_stat
```

Judging from Ha we see that this is a one-tailed test because alternative is that the difference is greater than zero i.e. right direction. Here can use direction = "right" or direction = "greater", they are equivalent.

```{r}
null_distribution %>%
  visualise() +
  shade_p_value(obs_stat = observed_stat, direction = "right")
```

```{r}
p_value <- null_distribution %>%
  get_p_value(obs_stat = observed_stat, direction = "right")

p_value
```
The p value is larger than our critical value of 0.05 and so we fail to reject H0 and conclude that we have not found enough evidence to suggest that omnivores sleep longer than herbivores on average. 
  

#3* Perform an appropriate statistical test to determine whether the proportion of domesticated animal types in the population of animal types is greater than $5\%$.
[**Hint** - think about creating an `is_domesticated` variable for the analysis]

```{r}
glimpse(msleep)
```

```{r}
msleep %>% 
  group_by(conservation) %>% 
    summarise()
```



```{r}
domesticated_animals <- msleep %>% 
  mutate(conservation = coalesce(conservation, "not_listed")) %>% 
  mutate(is_domesticated = ifelse(conservation == "domesticated", "domesticated", "not_domesticated"))

domesticated_animals
```

```{r}
domesticated_animals %>% 
  group_by(is_domesticated) %>% 
  count()
```
to determine whether the proportion of domesticated animal types in the population of animal types is greater than $5\%$with significance level α=0.05).
H0: πdomesticated - πnot_domesticated = 0
Ha: πdomesticated - πnot_domesticated ≠ 0
We treat is_domesticated as the response variable , with ‘domesticated’ as the ‘success’ category.


#had mind meltdown trying to work this out so moved onto interpretation
visualise data
```{r}
ggplot(domesticated_animals, aes(x = domesticated_animals, fill = )) +
  geom_bar(position = "fill") +
  coord_flip()
```


calculate observed statistic
```{r}
observed_stat <- domesticated_animals %>% 
  specify(is_domesticated, success = "domesticated") %>% 
  calculate(stat = "diff in props", order = c("domesticated", "not_domesticated"))

observed_stat


```


## Hypothesis testing - interpretation

### Defining the Hypothesis 
For the following three business problems 
-write out $H_0$ and $H_a$ in both mathematical notation and in words. -Also state the method you would use to generate the null distribution (bootstrap, permutation or simulation). 

1. You work for a independent coffee shop. You've performed a small survey in the local town and found that 40% of the people you randomly asked were aware of your shop in the town. You then conduct a marketing campaign by flyering local venues and targeting advertisements on social media. Finally you conduct a second small random survey in the town, asking if people are aware of your coffee shop. You want to test the hypothesis that the campaign has significantly increased awareness of the shop. 


##
H0: S2_Campaign awareness is the same as S1_campaign_awareness.
H1: S2_campaign_awareness is greater than S1_capmaign_awareness

H0: μcampaign_awareness(S1)−μcampaign_awareness(S2)=0
Ha: μcampaign-awareness(S1)−μcampaign_awareness(S2)>0

bootstrap



2. You work for a website design company and have performed an **A/B test** on the position of a banner on a website page promoting a particular item. 
<br><br>
<div class='emphasis'>
**A/B testing** 
A method comparing two versions of a web page, email, flyer or other marketing device against each other to determine which version performs better. As it is essentially a controlled experiment, the design should try to ensure that the groups experiencing both versions of the marketing device are need to establish that the two groups are equivalent and representative of the population.
In the current test, the first group continues to be shown the banner at the right hand side of the webpage (its usual position) while the test group is shown it at the top of the page. The performance metric we will be testing is click through rate (CTR) on the banner, i.e. what proportion of users click on the banner 

3. You work as an analyst for a car manufacturing company - they have specific standards they must meet for standards and regulation purposes. You have been asked to check the quality control of the manufacture of a particular car part. The part must have a width of 145mm, with a small (given) level of tolerance. You have been given data on a sample of $1,000$ parts produced over the period of a week. 

### Interpreting the results
For the 3 business problems stated above, imagine we performed you got the following p-values (with the given significance levels) write out your interpretation of the results. 

1. **Coffee shop problem**. Significance level: 0.05, calculated $p$-value: 0.07
p-value is greater than α so we lack sufficient evidence to reject H0 and so we fail to reject H0. Based on our sample, we do not have enough evidence that the new campaign has effected any positive increase in market awareness.

2. **Website company problem**. Significance level: 0.01, $p$-value: 0.006
p-value is less than α so reject H0 in favour of Ha. There is enough evidence in the sample to suggest that the sample is statistically significantly different from the null value.

3. **Manufacturing company problem**. Significance level: 0.05, $p$-value: 0.55
p-value is greater than α so lack sufficient evidence to reject H0 and fail to reject H0. Wwe do not have enough evidence that the new campaign has effected any positive increase in market awareness.
