---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
library( )
```

```{r}
library(stringr)
```


1. Load the `code_clan_tweets.csv` data. 
Find the number of rows, columns, and list all the variable names.
```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")
  nrow(code_clan_tweets)
  
```

 
```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")
ncol(code_clan_tweets)
```
 
```{r}
ls(code_clan_tweets)
```
```{r}
view(code_clan_tweets)
```
 

2. Find the total number of favourited tweets (stored in `favorite_count`) that CodeClan tweets have got. 
Don't add a tweet's `favorite_count` to the total when the tweet was quoted (see the `is_quote` variable). 
#this is number of rows not number of tweets

```{r}
number_favorited_tweets <- code_clan_tweets %>% 
  select(favorite_count) %>% 
filter(favorite_count != 0)
count(number_favorited_tweets)

```

need to select "favorite_count" and ! "is_quote" but this isn't right
```{r}

code_clan_tweets %>%
  filter(favorite_count & is_quote == FALSE) %>%
  tally(favorite_count)
```



3. Summarise the mean number of retweets (`retweet_count`), split by the type of platform that was used to tweet (`source`). 
Again, omit quoted tweets from the mean.

```{r}
              code_clan_tweets %>% 
                      filter(is_quote == FALSE) %>% 
                      group_by(source) %>% 
                      summarise(average_retweet = mean(retweet_count))
```


4. Count the total number of likes (i.e. total of `favorite_count`), split by `media` type, and arrange them from most likes to least. Assume that any tweets without a listed `media` type are of type "text". 
```{r}
total_likes <- code_clan_tweets %>% 
                 select(favorite_count, media_type) %>% 
                  group_by(media_type, "NA" == "text") %>% 
                  tally(favorite_count)

                  
total_likes
                    
```
4 answer:

```{r}
codeclan_tweets %>%
  mutate(media_type2 = replace_na(media_type, "text")) %>%
  group_by(media_type2) %>%
  summarise(favourite_count_total = sum(favorite_count)) %>%
  arrange(desc(favourite_count_total))
```



5. Find the mean number of characters that a CodeClan tweet contains.

#answers next day
```{r}
codeclan_tweets %>%
  summarise(avg_tweet_length = mean(display_text_width))
```


6. The `code_clan_info.csv` data file contains `status_url` along with other info. Load this in, and join it to the `code_clan_tweets` tibble, so that you have a `status_url` for each tweet. Decide which variable to join the tibbles on. 

#answers next day
```{r}
codeclan_info <- read_csv("data/code_clan_info.csv")
names(codeclan_info)
codeclan_all <- left_join(codeclan_tweets,codeclan_info, by = "tweet_id")
```


7. From your new joined data, create a new tibble `codeclan_hashtags` containing only the `tweet_id` and hashtags in lowercase for analysis. Keep only those tweets with hashtags.

#answers next day
```{r}
codeclan_hashtags <- codeclan_all%>% 
  select(tweet_id, hashtags) %>%
  mutate(lowcase_hashtag = str_to_lower(hashtags)) %>%
  select(-hashtags) %>%
  drop_na(lowcase_hashtag)

codeclan_hashtags
```

