---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
library( )
```

```{r}
library(stringr)
```


1. Load the `code_clan_tweets.csv` data. 
Find the number of rows, columns, and list all the variable names.
```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")
  nrow(code_clan_tweets)
  
```

 
```{r}
code_clan_tweets <- read_csv("data/code_clan_tweets.csv")
ncol(code_clan_tweets)
```
 
```{r}
ls(code_clan_tweets)
```
```{r}
view(code_clan_tweets)
```
 

2. Find the total number of favourited tweets (stored in `favorite_count`) that CodeClan tweets have got. 
Don't add a tweet's `favorite_count` to the total when the tweet was quoted (see the `is_quote` variable). 
#this is number of rows not number of tweets

```{r}
number_favorited_tweets <- code_clan_tweets %>% 
  select(favorite_count) %>% 
filter(favorite_count != 0)
count(number_favorited_tweets)

```

need to select "favorite_count" and ! "is_quote" but this isn't right
```{r}

code_clan_tweets %>%
  filter(favorite_count & is_quote == FALSE) %>%
  tally(favorite_count)
```



3. Summarise the mean number of retweets (`retweet_count`), split by the type of platform that was used to tweet (`source`). 
Again, omit quoted tweets from the mean.

```{r}
code_clan_tweets %>% 
                      filter(retweet_count & is_quote == FALSE) %>% 
                      group_by(source) %>% 
                      tally(mean_number_retweets)


```


4. Count the total number of likes (i.e. total of `favorite_count`), split by `media` type, and arrange them from most likes to least. Assume that any tweets without a listed `media` type are of type "text". 
```{r}
total_likes <- code_clan_tweets %>% 
                 filter(favorite_count) %>% 
                    filter(group_by(media_type))
                      arrange(desc(total_likes)
                    
```



5. Find the mean number of characters that a CodeClan tweet contains. 


6. The `code_clan_info.csv` data file contains `status_url` along with other info. Load this in, and join it to the `code_clan_tweets` tibble, so that you have a `status_url` for each tweet. Decide which variable to join the tibbles on. 


7. From your new joined data, create a new tibble `codeclan_hashtags` containing only the `tweet_id` and hashtags in lowercase for analysis. Keep only those tweets with hashtags.