---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
backpack
```
9 columns
```{r}
names(backpack)
```
#1_Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.

#2_What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.

#almost a normal distribution 

```{r}
ggplot(backpack) +
  aes(x = backpack_weight) +
  geom_bar(fill = "blue")
```


#3_Are there differences in backpack weight between sex?
#majority of women tend to carry  backpacks in weigh range 5-15kg
```{r}
ggplot(backpack) +
  aes(x = backpack_weight, fill = sex) +
  geom_bar()
```


#4_What is the relationship between body weight and backpack weight?
#as bodyweight increases  backpack weight has a slight positive correlation but not marked.
```{r}
ggplot(backpack) +
 aes(x = body_weight, y = backpack_weight) +
geom_point(colour = "blue")
```


#5_Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.
#different yes. Men tend to carry majority of backpacks in weight range 5-20kg
```{r}
ggplot(backpack) +
 aes(x = body_weight, y = backpack_weight, colour = sex) +
geom_point()
```

#6_What proportion of students report back pain?
#maybe this is unnecessary.

```{r}
backpain <- backpack %>% 
          select(back_problems, sex) %>% 
         summarise(back_problems) %>% 
          count(back_problems)
backpain

```


```{r}
ggplot(backpain) +
  aes(x = , fill = back_problems) +
  geom_bar(position = "fill")
```




Question 7.

Is this proportion different between men and women?
fill by gender
Yes.


```{r}
ggplot(backpack) +
  aes(x = back_problems, fill = sex,) +
  geom_bar(position = "fill")
```


Question 8.

Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.
```{r}
backpack
```

  hmm..trying to wrap my head around it
```{r}
ggplot(backpack)  +
  aes(x =body_weight, fill = sex) +
    geom_histogram(col = "white")
     # facet_wrap( body_weight ~ back_problems)
```



Try both backpack_weight and body_weight as well as the ratio.

Hint: faceting would be useful here.

```{r}
ggplot(data=dat, aes(x=Types, y=Number, fill=sample)) + 
  geom_bar(position = 'dodge', stat='identity') +
  geom_text(aes(label=Number), position=position_dodge(width=0.9), vjust=-0.25)
```

