---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
library(here)
```

```{r}
library(janitor)
```

```{r}
phones <- read_csv(here("data/20190928-items.csv")) %>% 
clean_names()


phones
```
```{r}
names(phones)
```
#1 Find the brand with the highest numbers of phones in the dataset.

#Samsung has 397 phones
```{r}
phones %>% 
  count(brand)

```


#2_For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution.
```{r}
samsung_ratings <- phones %>% 
  select(brand, rating) %>% 
  filter(brand == "Samsung")

samsung_ratings
```

```{r}
count(samsung_ratings)
```

```{r}
samsung_stats <- phones %>% 
  filter(brand == "Samsung") %>% 
  summarise(
    num =n(),
    mean = mean(rating),
    sd = sd(rating)
  )

samsung_stats
```

```{r}
phones %>% 
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = samsung_stats$mean ,
      sd =samsung_stats$sd
    ),
    col = "red"
  )
```



#2a_Do you think the normal distribution provides a good model of these ratings?
#looks quite good but could be slightly higher?



#3 We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the `mean()` and `sd()` of your distribution in an earlier question.

#Use these values to calculate the *proportions* of ratings within one-, two- and three standard deviations of the mean. 
#79% within 1sd
#92% within 2sd
#98% within 3sd
```{r}
samsung_ratings %>% 
  filter(rating >= samsung_stats$mean - samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + samsung_stats$sd) %>%
 # count() for sanity check 
  summarise(prop_within_1sd = n() / nrow(samsung_ratings))
  
```

```{r}
samsung_ratings %>% 
filter(rating >= samsung_stats$mean - (2*samsung_stats$sd)) %>%
filter(rating <= samsung_stats$mean + (2*samsung_stats$sd)) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_ratings))
```

```{r}

samsung_ratings %>% 
filter(rating >= samsung_stats$mean - (3*samsung_stats$sd)) %>%
filter(rating <= samsung_stats$mean + (3*samsung_stats$sd)) %>%
  summarise(prop_within_3sd = n() / nrow(samsung_ratings))
```


#Compare these proportions with those you expect for a normal distribution.  


```{r}

pnorm(1) - pnorm(-1)
pnorm(2) - pnorm(-2)
pnorm(3) - pnorm(-3)
```
#This samsung_ratings proportions are better fits for the normal distribution at 2sd from the mean than 1sd and even better at 3sd from the mean
