---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

```{r}
ames_house_data <- read_csv("data/ames.csv") %>% 
  clean_names()

ames_house_data
```


```{r}
glimpse(ames_house_data)
```

```{r}
ames_house_data %>% 
  select(lot_area)
```

#1_Investigate the distribution of `lot_area`. Is the distribution roughly normal? If not, what problems do you find?
#not really. majority of the data is concentrated around the mode.
```{r}
ames_lot_area_plot <- ames_house_data %>% 
  ggplot(aes(x = lot_area)) +
  geom_histogram()

ames_lot_area_plot
```

```{r}
summary_popn <- ames_house_data %>%
  summarise(
    mean_lot_area = mean(lot_area)
  ) 
summary_popn
```



* Compute and visualise a bootstrap sampling distribution for the `mean(lot_area)` of the sold houses.
```{r}
library(infer)
```

```{r}
infer_resample <- ames_house_data %>%
  specify(response = lot_area) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")

infer_resample%>%
  visualise(bins = 30)
  
```


# Use your bootstrap distribution to calculate a $95\%$ CI for `mean(lot_area)`, and visualise it on the distribution

```{r}
infer_ci_95 <- infer_resample %>%
  get_confidence_interval(level = 0.95, type = "percentile")

infer_ci_95
```

```{r}
infer_resample %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95)
```

* You would like to know the `mean(lot_area)` of the sold houses with higher confidence. Calculate the $99\%$ CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the $95\%$ CI? 
s#lightly wider. Yes. higher CI implies wider area 

```{r}
infer_ci_99 <- infer_resample %>%
  get_confidence_interval(level = 0.99, type = "percentile")

infer_ci_99
```

```{r}
infer_resample %>%
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_99)
```


# Calculate the point estimate of the `mean(lot_area)`

<hr>
```{r}
infer_resample %>%
  summarise(mean(stat))
```

