---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
```

```{r}
titanic <- read_csv("data/titanic.csv")
head(titanic)
```


Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)
```{r}
titanic %>% 
  distinct(class)
```

```{r}
crew_recode <- titanic %>% 
# filter(class) %>% 
   mutate(crew = recode(class,
                        "engineering crew" = TRUE,
                        "victualling crew" = TRUE,
                        "restaurant staff" = TRUE,
                        "deck crew" = TRUE,
                        .default = FALSE)) 
crew_recode
  
```
prob(crew)
```{r}
crew_recode %>% 
  count(crew == TRUE)
```
Therefore 2207 total complement.
Prob(crew) = 890 / (1317+890)
           = 890 / 2207
           =0.4032
        
```{r}
prob_crew <- crew_recode %>% 
  filter(crew == TRUE) %>% 
  summarise(prob =n()/ nrow(crew_recode))

prob_crew
```
      
      
```{r}
prob_survive_and_crew <- crew_recode %>% 
  filter(survived == "yes" & crew == TRUE ) %>% 
  summarise(prob =n()/ nrow(crew_recode))

  prob_survive_and_crew
```
confirming this 
prob(crew and survived) = 211/2207 = 0.095

```{r}

crew_recode %>% 
  count(crew == TRUE & survived == "yes")
```
prob(survived | crew = prob(survived AND crew) / prob(crew)
            
```{r}
prob_survive_given_crew = prob_survive_and_crew/prob_crew
prob_survive_given_crew

```
to confirm via dplyer


```{r}
prob_survived_given_crew = crew_recode %>% 
  filter(crew == TRUE) %>% 
  summarise(prob = sum(survived == "yes")/n())
prob_survived_given_crew
```
```{r}
prob_survived_given_passenger <- crew_recode %>% 
  filter(crew == FALSE) %>% 
  summarise(prob = sum(survived == "yes")/n())
prob_survived_given_passenger
```
#Therefore:
prob_survived_given_crew = 0.237 while
prob_survived_given_passenger = 0.380 (to 3sf)
Therefore it was more likely to survive given a passenger

#Part2
1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?

#prob(x < 5) = 4/6


2. How many outcomes are there in the sample space of an experiment in which we toss three dice?

# S = 3cubed = 27


3. In an experiment where we roll four dice:how many outcomes are in the event total_equals_5_6_7 (i.e. the total of the four dice equals 5, 6 or 7)?

Sample space = 6^4 = 1296. 
(total_equals_5_6_7)


#what is the probability of this event?
#what assumptions do we apply in answering part ii?
 Hint

4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

